---
title: "PCA_PC_PLS"
output: html_document
date: "2023-12-04"
---

```{r packages}
library(dplyr)    # for data manipulation
library(ggplot2)  # for awesome graphics

# Modeling packages
library(caret)    # for cross-validation, etc.

# Model interpretability packages
library(vip)      # variable importance
library(h2o)
```

```{r View data1}
#data1 <- read.csv("US-pumpkins.csv", header = TRUE)
data1 <- AmesHousing::make_ames()
#data1(Hitters)
#data1 <- Hitters
dim(data1) 
View(data1)
```

chr: Character vector, lgl:logical vector, dbl:Double(same as numeric) vector), 
```{r Structure data}
library(dplyr)    # for data manipulation
dim(data1)
glimpse(data1) #From tidyverse library
#summary(data1)
```
Requiremtns to perform PCA
1. Data is in tidy format per Wickham and others (2014);
2. Any missing values in the data must be removed or imputed;
3. Typically, the data must all be numeric values (e.g., one-hot, label, ordinal encoding categorical features);
4. Numeric data should be standardized (e.g., centered and scaled) to make features comparable.

Set a seed and randomly partition the data into training and test set (70%/30%)
```{r Datasplit train test}
library(rsample)
set.seed(123)
split <- initial_split(data1, 
                       prop = 0.7, 
                       strata = "Sale_Price")
data_train  <- training(split)
data_test   <- testing(split)

x_train <- model.matrix(Sale_Price ~ ., data = data_train)[, -1]
y_train <- data_train$Sale_Price

x_test <- model.matrix(Sale_Price ~ ., data = data_test)[, -1]
y_test <- data_test$Sale_Price
```

```{r PCR Regression}
library(caret)
set.seed(123)
cv_model_pcr <- train(
  Sale_Price ~ ., 
  data = data_train, 
  method = "pcr", # see method here
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"), # an integrated method for basic data preprocessing (preProcess()) within caret
  tuneLength = 100 # tuning parameter regarding quantity of principal components
)
cv_model_pcr$results %>% # model with lowest RMSE
  dplyr::filter(ncomp == pull(cv_model_pcr$bestTune))
plot(cv_model_pcr$results$RMSE)
```

```{r PLS Regression}
set.seed(123)
cv_model_pls <- train(
  Sale_Price ~ ., 
  data = data_train, 
  method = "pls",
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"),
  tuneLength = 30 #using less PC in PLS
)
cv_model_pls$results %>%
  dplyr::filter(ncomp == pull(cv_model_pls$bestTune))
plot(cv_model_pls$results$RMSE)
```

```{r PLS Variable importance }
library(vip)      # variable importance
vip(cv_model_pls, num_features = 20, method = "model")
#  from 100 (most important) to 0 (least important)
```

```{r Proportion of variance}
library(dplyr)
data1_numeric <- data_train %>% 
  mutate(across(everything(), as.numeric))

pca_out = prcomp(data1_numeric, scale=TRUE)
summary(pca_out)
```
